---
title: "Touch Paper"
output: pdf_document
---

The analysis we did of the stats behind figure 1 was sub-standard. The model fitted should be a two-factor linear mixed effect model (Genotype * treatment are the two factors, w/ interaction), with a random effect term for the tray (batch) effect. This removes the error in each of the two model factors caused by tray effect whilst fitting the model.

So, lets get the data:
```{r read_data}
fig1_data <- read.csv("TouchData.csv")
fig1_data$Genotype <- relevel(fig1_data$Genotype, ref="WT")
summary(fig1_data)
```

Now, for each phenotype, we want to fit the linear mixed effects model. We'll need the package `nlme`, which contains the `lme` function we use below.


```{r lme_models} 
library(nlme)

phenotypes <- unique(as.character(fig1_data$Measurement))
for (pheno in phenotypes) {
  pdata <- fig1_data[fig1_data$Measurement == pheno,]
  fit <- lme(Value ~ Genotype * Treatment, data = pdata, random =~ 1|Tray)
  print(pheno)
  print(summary(fit))
#   g.fit <- tapply(predict(fit),paste(gp$Genotype,gp$Treatment),mean)
#   oo <-  barplot(g.fit,ylim = c(0,20),main = "PetLength")
#   segments(oo,g.fit,oo,g.fit+fit$sigma)
#   segments(oo,g.fit,oo,g.fit-fit$sigma)
}
```

So, we have the following "ANOVA" table (though this is NOT an anova). The values are the size of the effect (in mm) and the p-value assoicated with the effect. This is all relative to non-touched wild-type plants.  

| ----------- | ----------- | ------------- | --------------------- |
| Phenotype   | Genotype    | Treatment     | Genotype:Treatment    |
| ----------- | ----------- | ------------- | --------------------- |
| Leaf Len    | -6.1 (0.00) | -1.3 (0.09)   | 3.4 (0.00)            |
| Leaf Width  | -3.4 (0.00) | -1.4 (0.00)   | 1.8 (0.00)            |
| Petiole Len | -3.7 (0.00) | -4.6 (0.00)   | 2.0 (0.00)            |


Which is pretty sweet.

What does this mean?
--------------------

Basically, this fits a linear model. If we go back to old school regression,
it's like fitting y = ax + b type formulae over the model "Phenotype depends on
Genotype + Treatement + an interaction between the two + a random effect from
the tray".  So, in the above table, the effects (relative to WT Non-touched)
indicated a model estimation of roughly how large the effect of genotype,
treatment, and the interaction is the size and direction of the interaction.
Positive interactions bascially mean that the effect is exaggerated in the
interaction, i.e. in this case, for Leaf Len, both genotype and treatment have
negative effects (SDG8 reduces leaf length, as does touching). Therefore, SDG8
touched plants have shorter leaves than just the addition of the effects of
genotype and treatment.


OK, now some plots. These are NOT actual values. 

```{r model_fit_plots, fig.width=6} 
for (pheno in phenotypes) {
  pdata <- fig1_data[fig1_data$Measurement == pheno,]
  fit <- lme(Value ~ Genotype * Treatment, data = pdata, random =~ 1|Tray)
  fit.data <- tapply(predict(fit),paste(pdata$Genotype,pdata$Treatment),mean)
  xx <-  barplot(fit.data,ylim = c(0,26),main = pheno)
  segments(xx,fit.data,xx,fit.data+fit$sigma)
  segments(xx,fit.data,xx,fit.data-fit$sigma)
  print(paste("Predicted values for", pheno))
}
```
