---
title: "Cazzonelli et al 2014 Figure 1"
output: pdf_document
---

The model fitted should be a two-factor linear mixed effect model
(Genotype * treatment are the two factors, w/ interaction), with a random
effect term for the tray (batch) effect. This removes the error in each of the
two model factors caused by tray effect whilst fitting the model.

So, lets get the data:
```{r read_data}
fig1_data <- read.csv("TouchData.csv")
fig1_data$Genotype <- relevel(fig1_data$Genotype, ref="WT")
fig1_data$Block <- as.numeric(fig1_data$Genotype)
fig1_data$Block12 <- as.numeric(fig1_data$Genotype) * as.numeric(fig1_data$Tray)
summary(fig1_data)
table(paste(fig1_data$Genotype, fig1_data$Treatment))
# this is your N per treatment
table(paste(fig1_data$Block, fig1_data$Tray, fig1_data$Measurement))
min(table(paste(fig1_data$Block, fig1_data$Tray, fig1_data$Measurement)))
max(table(paste(fig1_data$Block, fig1_data$Tray, fig1_data$Measurement)))
```

Now, for each phenotype, we want to fit the linear mixed effects model. We'll
need the package `nlme`, which contains the `lme` function we use below.


```{r lme_models} 
library(nlme)

phenotypes <- unique(as.character(fig1_data$Measurement))
for (pheno in phenotypes) {
  pdata <- fig1_data[fig1_data$Measurement == pheno,]
  fit <- lme(Value ~ Genotype * Treatment, data = pdata, random =~ 1|Tray)
  print(pheno)
  print(summary(fit))
}
```

So, we have the following "ANOVA" table (though this is NOT an anova). The
values are the size of the effect (in mm) and the p-value assoicated with the
effect. This is all relative to non-touched wild-type plants.  

| ----------- | ----------- | ------------- | --------------------- |
| Phenotype   | Genotype    | Treatment     | Genotype:Treatment    |
| ----------- | ----------- | ------------- | --------------------- |
| Leaf Len    | -6.2 (0.00) | -1.3 (0.24)   | 3.5 (0.00)            |
| Leaf Width  | -3.4 (0.00) | -1.4 (0.01)   | 1.8 (0.00)            |
| Petiole Len | -3.7 (0.00) | -4.6 (0.01)   | 2.0 (0.01)            |


What does this mean?
--------------------

This fits the model "Phenotype is explained by Genotype + Treatement + an
interaction between the two + a random effect from the tray".  So, in the above
table, the effects (relative to WT Non-touched) indicate a model estimation of
roughly how large the effect of genotype, treatment, and the interaction is the
size and direction of the interaction. In this case, for Leaf Len, both genotype
and treatment have negative effects (SDG8 reduces leaf length, as does
touching). Therefore, SDG8 touched plants have longer (more positive) leaves
than just the addition of the effects of genotype and treatment.

Plots
=====

OK, now some plots. These are NOT actual values, they are fitted values, based
off the same model we used above. This means that these values don't nicely
correspond to any exact value in a table of raw data, but means that they are
more accurately annotated with stats from the model above.

```{r model_fit_plots, fig.width=6} 
for (pheno in phenotypes) {
  pdata <- fig1_data[fig1_data$Measurement == pheno,]
  N = nrow(pdata)
  fit <- lme(Value ~ Treatment * Genotype, data = pdata, random =~ 1|Tray)
  fit.data <- tapply(predict(fit),
                     paste(pdata$Genotype, pdata$Treatment, sep='\n'),
                     mean)
  # re-order the factors so it's WT then ccr2
  fit.data <- fit.data[c(3,4,1,2)]
  names(fit.data) <- c("WT\nCtrl", "WT\nTch", "sdg8\nCtrl", "sdg8\nTch")
  # plot to screen
  xx <-  barplot(fit.data, ylim = c(0,25), beside = T,
                 col=c("#000000", "#FFFFFF", "#000000", "#FFFFFF"))
  se = fit$sigma / sqrt(N)
  segments(xx, fit.data, xx, fit.data + (2 * se))
  segments(xx, fit.data, xx, fit.data - (2 * se))
  title(paste("LME-predicted values for", pheno))
  # re-plot to pdf
  pdf(paste0(pheno, '_fitted.pdf'), width=3.5)
  xx <-  barplot(fit.data, ylim = c(0,25), beside = T,
                 col=c("#000000", "#FFFFFF", "#000000", "#FFFFFF"))
  se = fit$sigma / sqrt(N)
  segments(xx, fit.data, xx, fit.data + (2 * se))
  segments(xx, fit.data, xx, fit.data - (2 * se))
  title(paste(pheno))
  dev.off()
  #
  print(paste("Predicted values for", pheno))
}
```
