Nazia's touch data
==================

Read in the data, and split out tables by measurement
```{r}
library(ggplot2)
library(plyr)

a <- read.csv("TouchData.csv", header=T, quote='')
summary(a)
str(a)


ll<-a[a$Measurement=='LeafLength',-5]
lw<-a[a$Measurement=='LeafWidth',-5]
pl<-a[a$Measurement=='PetioleLength',-5]
summary(ll)
summary(lw)
summary(pl)
```


Then, do boxplots, AOV and Tukey's HSD on the full dataset, and each of the split datasets.
```{r}
boxplot(Value~Genotype+Treatment,data=pl)
a_aov<-aov(Value~Genotype*Treatment+Measurement,data=a)
summary(a_aov)
TukeyHSD(a_aov)
```

Leaf Length shows significant Geno effect & Geno:Treatement interaction effect. There are
significant differerences between Genotype & Treatement points as shown in the Tukey HSD output
under $Genotype:Treatment.
```{r}
ll_aov<-aov(Value~Genotype*Treatment,data=ll)
summary(ll_aov)
TukeyHSD(ll_aov)
```

Leaf width show very significant Geno effect, sig. treatment effect, & a sig. interaction term.
Tukeys HSD shows the only non-signficant comparison is *ccr1.1* Touch vs Non-touched.
```{r}
lw_aov<-aov(Value~Genotype*Treatment,data=lw)
summary(lw_aov)
TukeyHSD(lw_aov)
```

Again, for petiole length, pretty much everything is significant. Only WT touched vs *ccr1.1* not touched is non-significant.
```{r}
pl_aov<-aov(Value~Genotype*Treatment,data=pl)
summary(pl_aov)
TukeyHSD(pl_aov)
```


Summarise and plot data in a pretty ggplot graph
```{r}
plotData <- ddply(a, c("Genotype","Treatment","Measurement"), summarise,
              mean = mean(Value, na.rm=TRUE),
              sd = sd(Value,na.rm=TRUE),
              n = sum(!is.na(Value)),
              se = sd/sqrt(n))

ggplot(plotData, aes(x=Genotype, y=mean,fill=Treatment )) +
  geom_bar(stat="identity", position="dodge") +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(width = 0.90)) +
  facet_wrap(~Measurement) +
  ylab("Length/Width (mm)")
```